Supplemental Material for: The evolution of gene expression in the term placenta of viviparous mammals

### 2015-11-20 manuscript/supplemental_material.Rnw
include list of tables and figures in supplemental material
switch to cleveref
abstract out utility functions and table code

\documentclass[english,12pt]{article}
\usepackage{fontspec}
\setmainfont{FreeSerif}
\setsansfont{FreeSans}
\setmonofont{FreeMono}
\usepackage[letterpaper,left=1in,right=1in,top=1in,bottom=1in,headheight=14pt]{geometry}
\usepackage{url}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage[bf]{caption}
\usepackage{rotating}
\usepackage{wrapfig}
\usepackage{booktabs}
\usepackage{float}
\usepackage{multirow}
\usepackage{setspace}
\usepackage{wrapfig}
% \usepackage{txfonts}
\usepackage[plain]{fancyref}
\renewcommand*{\fref}[1]{\Fref{#1}}
\usepackage{acronym}
\usepackage{array}
\usepackage{dcolumn}
\usepackage{adjustbox}
\usepackage{longtable}
\usepackage{pdflscape}
\usepackage[backend=biber,natbib=true,hyperref=true,style=reading,autocite=inline]{biblatex}
\addbibresource{references.bib}
\usepackage[noblocks,auth-sc]{authblk}
\usepackage[hyperfigures,bookmarks,colorlinks]{hyperref}
%\usepackage[hyperfigures,bookmarks,colorlinks,citecolor=black,filecolor=black,linkcolor=black,urlcolor=black]{hyperref}
% Floats at end for journals
% \usepackage[nolists,figuresfirst,nomarkers,heads,nofighead,notabhead]{endfloat}
% \AtBeginFigures{\section*{Figure Legends}\renewcommand{\efloatseparator}{\hbox{}}}
% \AtBeginTables{\section*{Tables}\renewcommand{\efloatseparator}{\mbox{}}}
\usepackage[nomargin,inline,draft]{fixme}
\newcommand{\DLA}[1]{\textcolor{red}{\fxnote{DLA: #1}}}
\usepackage[x11names,svgnames]{xcolor}
\usepackage{texshade}
\usepackage{tikz}
\usepackage{nameref}
\usepackage{zref-xr,zref-user}
\newenvironment{narrow}[2]{%
  \begin{list}{}{%
      \setlength{\topsep}{0pt}%
      \setlength{\leftmargin}{#1}%
      \setlength{\rightmargin}{#2}%
      \setlength{\listparindent}{\parindent}%
      \setlength{\itemindent}{\parindent}%
      \setlength{\parsep}{\parskip}}%
  \item[]}{\end{list}}
\def\newblock{\hskip}
\newenvironment{paperquote}{%
  \begin{quote}%
     \it
  }%
  {\end{quote}}
\renewcommand{\textfraction}{0.15}
\renewcommand{\topfraction}{0.85}
\renewcommand{\bottomfraction}{0.65}
\renewcommand{\floatpagefraction}{0.60}
%\renewcommand{\baselinestretch}{1.8}
\newenvironment{enumerate*}%
  {\begin{enumerate}%
    \setlength{\itemsep}{0pt}%
    \setlength{\parskip}{0pt}}%
  {\end{enumerate}}
\newenvironment{itemize*}%
  {\begin{itemize}%
    \setlength{\itemsep}{0pt}%
    \setlength{\parskip}{0pt}}%
  {\end{itemize}}
\oddsidemargin 0.0in 
\textwidth 6.5in
\raggedbottom
\clubpenalty = 10000
\widowpenalty = 10000
% make the afill smaller
\renewcommand\Affilfont{\itshape\small}
\renewcommand\cite[1]{\autocite{#1}}
\pagestyle{fancy}

@@

@@

capwords <- function(s, strict = FALSE) {
    cap <- function(s) paste(toupper(substring(s, 1, 1)),
                             {s <- substring(s, 2); if(strict) tolower(s) else s},
                             sep = "", collapse = " " )
    sapply(strsplit(s, split = " "), cap, USE.NAMES = !is.null(names(s)))
}

capfirst <- function(s, strict = FALSE) {
    paste(toupper(substring(s, 1, 1)),
          {s <- substring(s, 2); if(strict) tolower(s) else s},
          sep = "")
}

species <- function(s,join=TRUE) {
    temp <- paste0("\\textit{",capfirst(s),"}")
    if (join) {
        return(english.join(temp))
    } else {
        return(temp)
    }
}

english.join <- function(s){
    ret <- c()
    ## everything but the last gets a , appended if there are more than two
    if (length(s) > 2) {
        s[-length(s)] <- 
            paste0(s[-length(s)],",")
    }
    ## the second to last gets an and appended if there are more than one
    if (length(s) > 1) {
        s[length(s)-1] <- 
            paste0(s[length(s)-1]," and")
    }
    return(paste(sep="",collapse=" ",s))
}

@@

@@

@@

<<human.expressed.genes,echo=FALSE,results="asis">>=
print(xtable(genes.expression.table(species.name="homo sapiens"),
             caption="Expression of all genes with FPKM $\\ge 50$ in \\textit{Homo sapiens}",
             label="tab:all_human_placenta"
             ),
      booktabs=TRUE,
      tabular.environment="longtable",
      floating=FALSE,
      caption.placement="top",
      add.to.row = list(pos = list(0),command = "\\midrule \\endhead "),
      size="\\tiny"
      )
@

>>>>

\documentclass[english,12pt]{article}
\usepackage{fontspec}
\setmainfont{FreeSerif}
\setsansfont{FreeSans}
\setmonofont{FreeMono}
\usepackage[letterpaper,left=1in,right=1in,top=1in,bottom=1in,headheight=14pt]{geometry}
\usepackage{url}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage[bf]{caption}
\usepackage{rotating}
\usepackage{wrapfig}
\usepackage{booktabs}
\usepackage{float}
\usepackage{multirow}
\usepackage{setspace}
\usepackage{wrapfig}
% \usepackage{txfonts}
\usepackage{acronym}
\usepackage{array}
\usepackage{dcolumn}
\usepackage{adjustbox}
\usepackage{longtable}
\usepackage{pdflscape}
\usepackage[backend=biber,natbib=true,hyperref=true,style=reading,autocite=inline]{biblatex}
\addbibresource{references.bib}
\usepackage[noblocks,auth-sc]{authblk}
\usepackage[hyperfigures,bookmarks,colorlinks]{hyperref}
%\usepackage[hyperfigures,bookmarks,colorlinks,citecolor=black,filecolor=black,linkcolor=black,urlcolor=black]{hyperref}
% Floats at end for journals
% \usepackage[nolists,figuresfirst,nomarkers,heads,nofighead,notabhead]{endfloat}
% \AtBeginFigures{\section*{Figure Legends}\renewcommand{\efloatseparator}{\hbox{}}}
% \AtBeginTables{\section*{Tables}\renewcommand{\efloatseparator}{\mbox{}}}
\usepackage[nomargin,inline,draft]{fixme}
\newcommand{\DLA}[1]{\textcolor{red}{\fxnote{DLA: #1}}}
\usepackage[x11names,svgnames]{xcolor}
\usepackage{texshade}
\usepackage{tikz}
\usepackage{nameref}
\usepackage{zref-xr,zref-user}
\usepackage{cleveref}
\newenvironment{narrow}[2]{%
  \begin{list}{}{%
      \setlength{\topsep}{0pt}%
      \setlength{\leftmargin}{#1}%
      \setlength{\rightmargin}{#2}%
      \setlength{\listparindent}{\parindent}%
      \setlength{\itemindent}{\parindent}%
      \setlength{\parsep}{\parskip}}%
  \item[]}{\end{list}}
\def\newblock{\hskip}
\newenvironment{paperquote}{%
  \begin{quote}%
     \it
  }%
  {\end{quote}}
\renewcommand{\textfraction}{0.15}
\renewcommand{\topfraction}{0.85}
\renewcommand{\bottomfraction}{0.65}
\renewcommand{\floatpagefraction}{0.60}
%\renewcommand{\baselinestretch}{1.8}
\newenvironment{enumerate*}%
  {\begin{enumerate}%
    \setlength{\itemsep}{0pt}%
    \setlength{\parskip}{0pt}}%
  {\end{enumerate}}
\newenvironment{itemize*}%
  {\begin{itemize}%
    \setlength{\itemsep}{0pt}%
    \setlength{\parskip}{0pt}}%
  {\end{itemize}}
\oddsidemargin 0.0in 
\textwidth 6.5in
\raggedbottom
\clubpenalty = 10000
\widowpenalty = 10000
% make the afill smaller
\renewcommand\Affilfont{\itshape\small}
\renewcommand\cite[1]{\autocite{#1}}
\pagestyle{fancy}

@@

\listoftables

\listoffigures

@@

@@

<<load.go.results,echo=FALSE,results="hide",cache.extra=tools::md5sum("../go_annotation/go_tissue_analysis")>>=
load("../go_annotation/go_tissue_analysis")

\section{Transcriptome statistics}

@@

\section{Gene Ontology Analyses}

\subsection{Top 100 Transcripts Enrichment}

<<go.analysis.all.species,echo=FALSE,results="asis">>=
for (this.species in names(species.names)) {
    cat(paste0("\\subsubsection{Signficantly enriched ($\\mathrm{FDR}\\le0.05$) GO terms in ",
                           " the top 100 \\textit{",
                           capfirst(this.species),"} expressed transcripts}\n\n"))
    this.caption <- paste0("Signficantly enriched ($\\mathrm{FDR}\\le0.05$) GO terms in ",
                           " the top 100 \\textit{",capfirst(this.species),"} expressed transcripts")
    this.label <- paste0("tab:go_terms_enriched_",gsub(" ","_",this.species))
    print(xtable(go.results[grepl("100 No Housekeeping",type) & FDR <= 0.05 & species == this.species,
                            list(GO.ID,Term,FDR,Ontology)
                            ][order(FDR)],
                 caption=this.caption,
                 label=this.label,
                 display=c("s","s","s","e","s")
                 ),
          size="\\small",
          booktabs=TRUE,
          tabular.environment="longtable",
          caption.placement="top",
          floating=FALSE)
}
@

\subsection{Top 5\% Transcripts Enrichment}

<<go.analysis.top.5p.all.species,echo=FALSE,results="asis">>=
for (this.species in names(species.names)) {
    cat(paste0("\\subsubsection{Signficantly enriched ($\\mathrm{FDR}\\le0.05$) GO terms in ",
                           " the top 5\\% of \\textit{",
                           capfirst(this.species),"} expressed transcripts}\n\n"))
    this.caption <- paste0("Signficantly enriched ($\\mathrm{FDR}\\le0.05$) GO terms in ",
                           " the top 5\\% of \\textit{",
                           capfirst(this.species),"} expressed transcripts.")
    print(xtable(go.results[grepl("Top 5% No Housekeeping",type) & FDR <= 0.05 & species == this.species,
                            list(GO.ID,Term,FDR,Ontology)
                            ][order(FDR)],
                 caption=this.caption,
                 label=paste0("tab:go_terms_enriched_top_five_",gsub(" ","_",this.species)),
                 display=c("s","s","s","e","s")
                 ),
          size="\\small",
          booktabs=TRUE, 
          tabular.environment="longtable",
          caption.placement="top",
          floating=FALSE)
}
@

@@

\section{Gene Expression}
<<top.expressed.genes,echo=FALSE,results="asis">>=
for (this.species in names(species.names)) {
    this.caption <- paste0("Expression of all genes with FPKM $\\ge 50$ in \\textit{",
                           capfirst(this.species),"}")
    cat(paste0("\\subsection{",this.caption,"}\n\n"))
    print(xtable(genes.expression.table(species.name=this.species),
                 caption=this.caption,
                 label=paste0("tab:all_",gsub(" ","_",this.species),"_placenta")
                 ),
          booktabs=TRUE,
          tabular.environment="longtable",
          floating=FALSE,
          caption.placement="top",
          size="\\tiny"
          )
}
@
